RegressionPvNP

### 2017-06-02 RegressionPvNP.Rmd
…

```{r setup, include=FALSE}
library(ggplot2)
library(ggfortify)
library(plyr)
library(pander)
library(reshape2)
library(gridExtra)
library(GenomicRanges)
library(glmnet)

@@

```{r tile, echo=F, message=F, warning=F, include=F}
tile.w=1e07
#tile.w=5e06
tile.files = grep('tile_.*1e\\+07', analysis.files, value=T)
#tile.files = grep('tile_.*5e\\+06', analysis.files, value=T)
load(tile.files[1], verbose=T)
dim(mergedDf)

#cache.dir = paste(data, 'Analysis/5e06', sep='/')
cache.dir = paste(data, 'Analysis/1e07', sep='/')
if (!dir.exists(cache.dir)) dir.create(cache.dir)

@@

  if (select == 'classification') {
    df$lambda.min = df$mean == max(df$mean)
    se1 = df[df$lambda.min == T, 'mean'] - df[df$lambda.min == T, 'sme']

@@

plot.patient.cv<-function(df, fit=NULL) {
  gp = ggplot(df, aes(y=mean,x=log.lambda)) + geom_point() + geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + 
    geom_point(data=subset(df, lambda.min == T), aes(y=mean, x=log.lambda, colour="lambda.min"), size=2 ) +  
    geom_vline(xintercept = subset(df, lambda.min == T)$log.lambda, colour='grey') +
    annotate("text", x=subset(df, lambda.min == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
             label=round(df[subset(df, lambda.min == T)$lambda.at,'mean'],3))

  gpD = ggplot(df, aes(y=mean.b.dev,x=log.lambda)) + geom_point() +
    geom_errorbar(aes(ymin=mean.b.dev-sme.b.dev, ymax=mean.b.dev+sme.b.dev)) + 
    geom_point(data=subset(df, lambda.min == T), aes(y=mean.b.dev, x=log.lambda, colour="lambda.min"), size=2 ) +          
    geom_vline(xintercept = subset(df, lambda.min == T)$log.lambda, colour='grey') +
    annotate("text", x=subset(df, lambda.min == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
             label=round(df[subset(df, lambda.min == T)$lambda.at,'mean'],3))

  if (nrow(subset(df, lambda.1se == T)) > 0) {
    gp = gp + geom_point(data=subset(df, lambda.1se == T), aes(y=mean, x=log.lambda, colour="lambda.1se"), size=2 ) +
      geom_vline(xintercept = subset(df, lambda.1se == T)$log.lambda, colour='grey') +
      annotate("text", x=subset(df, lambda.1se == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
               label=round(df[subset(df, lambda.1se == T)$lambda.at,'mean'],3))

    gpD = gpD + geom_point(data=subset(df, lambda.1se == T), aes(y=mean.b.dev, x=log.lambda, colour="lambda.1se"), size=2 ) +
      geom_vline(xintercept = subset(df, lambda.1se == T)$log.lambda, colour='grey') +
      annotate("text", x=subset(df, lambda.1se == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
               label=round(df[subset(df, lambda.1se == T)$lambda.at,'mean'],3)) 
  }

@@

```{r xvalpt, message=F, warning=F, echo=T, fig.height=16, fig.width=16}
nl = 3000

@@

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  for (a in alpha.values) {
    fit0 <- glmnet(dysplasia.df, labels, alpha=a, nlambda=nl, family='binomial') # all patients
    #autoplot(fit0, xvar='lambda', main=paste('fit0, all samples, alpha=',a,sep='')) + theme(legend.position='none')

    cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=fit0$lambda, pts=pts, a=a, nfolds=folds, splits=splits, fit=fit0)

@@

dysplasia.df[ labels == 1 ,regions]

dysplasia.df[ labels == 0 ,regions]

@@

```{r, warning=F, message=F, eval=F}
#topF = do.call(rbind, strsplit( names(which((coef.stable[['1']]) >= 0.75)), ':|-'))
#colnames(topF) = c('chr','start','end')
#topF = apply(topF, 2, as.numeric)

topF = names(which((coef.stable[['1']]) >= 0.75))

library(biomaRt)
mart = biomaRt::useMart("ENSEMBL_MART_ENSEMBL", host="grch37.ensembl.org")
biomart = useDataset("hsapiens_gene_ensembl", mart)

attr = c('ensembl_gene_id','hgnc_symbol','start_position','end_position', 'chromosome_name', 'go_id', 'entrezgene')
bm = getBM(mart=biomart, attributes=attr, filters=list('chromosomal_region' = topF))
bm = bm[-which(bm$hgnc_symbol == ""),] 
#head(bm)

bm2 = ddply(bm, .(entrezgene, ensembl_gene_id, hgnc_symbol), summarise, 'chrom'=unique(chromosome_name), 'start'=min(start_position), 'end'=max(end_position), 'GO.IDs'=paste(go_id, collapse=','))

ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)

bm2 = subset(bm2, hgnc_symbol %in% ccgenes$Gene.Symbol)

features.in.prog = dysplasia.df[names(which(labels == 1)),topF]

#features.in.prog <= 0.9
#features.in.prog >= 1.1

@@

@@

file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  # No HGD/IMC
  samples = intersect(rownames(dysplasia.df), subset(info, Pathology %nin% c('HGD', 'IMC'))$Samplename)
  for (a in alpha.values) {
    # all patients
    fitNoHGD <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=nl)

    cv.nohgd = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitNoHGD$lambda, pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fitNoHGD)

@@

# Leave one out

@@

    # Fit generated on all samples, including HGD
    a = 1
    fitLOO <- glmnet(training, labels[train.rows], alpha=a, family='binomial', nlambda=nl) # all patients
    cv = crossvalidate.by.patient(x=training, y=labels[train.rows], lambda=fitLOO$lambda, 
                                  pts=subset(pts, Samplename %in% samples), a=a, nfolds=10, splits=5, fit=fitLOO)
    #plots[[pt]] = cv$plot + ggtitle(pt)

@@

back = lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(pt) {
  pt = pt %>% group_by(PID, Endoscopy.Year) %>% summarise(
                  'pred'=sum(as.integer(Prediction > 0.5)), 'samples'=length(PID), 
                  'final.endo'= length(which(grepl('HGD|IMC', Pathology))) > 0,
                  'months.before.final'= (max(pt$Endoscopy.Year) - unique(Endoscopy.Year))*12 )
  arrange(pt, -Endoscopy.Year, PID)
})

endos = do.call(rbind, lapply(1:max(sapply(back,nrow)), function(i) {
  df = do.call(rbind, lapply(back, function(x) {
    if (i > nrow(x)) return(NA)
    cbind(x[i, c('Endoscopy.Year','pred','samples','final.endo', 'months.before.final')], i)
  }))
  df$Patient = rownames(df)
  return(df)
}))

@@

```{r, echo=F, warning=F, message=F, fig.height=8, fig.width=8}
plot.endo( subset(endos, Patient %in% subset(me, max.endo <= 2 & max.endo > 0)$Patient), min(endos$months.before.final), max(endos$months.before.final)  )
```

```{r, echo=F, warning=F, message=F, fig.height=4, fig.width=4}
plot.endo( subset(endos, Patient %in% subset(me, max.endo == 0)$Patient),min(endos$months.before.final), max(endos$months.before.final) )
```

@@

binomial.deviance

@@

# Randomize the labels

@@

>>>>

```{r setup, include=FALSE}
library(biomaRt)
library(ggplot2)
library(ggfortify)
library(GGally)
library(plyr)
library(pander)
library(reshape2)
library(gridExtra)
library(GenomicRanges)
library(glmnet)

@@

```{r tile, echo=F, message=F, warning=F, include=F}
tile.w=1e07
#tile.w=5e06
tile.files = grep(as.character(tile.w), analysis.files, value=T, fixed=T)
load(tile.files[1], verbose=T)
dim(mergedDf)

cache.dir = paste(data, 'Analysis',sub('\\+', '', tile.w), sep='/')
if (!dir.exists(cache.dir)) dir.create(cache.dir)

@@

  if (select == 'classification') {
    # Minimize the classification mean error & select min lambda
    df[which.max(subset(df, sme < median(sme) & sme > min(sme))$mean), 'lambda.min'] = T

    #df$lambda.min = df$mean == max(df$mean)
    se1 = df[df$lambda.min == T, 'mean'] - df[df$lambda.min == T, 'sme']

@@

plot.patient.cv<-function(df, fit=NULL) {
  gp = ggplot(df, aes(y=mean,x=log.lambda)) + geom_point() + geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + 
    geom_point(data=subset(df, lambda.min == T), aes(y=mean, x=log.lambda, colour="lambda.min"), size=2 ) +  
    geom_vline(xintercept = subset(df, lambda.min == T)$log.lambda, colour='grey') +
    annotate("text", x=subset(df, lambda.min == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
             label=paste(round(df[subset(df, lambda.min == T)$lambda.at,c('mean', 'sme')],3), collapse='\n+/-'))

  gpD = ggplot(df, aes(y=mean.b.dev,x=log.lambda)) + geom_point() +
    geom_errorbar(aes(ymin=mean.b.dev-sme.b.dev, ymax=mean.b.dev+sme.b.dev)) + 
    geom_point(data=subset(df, lambda.min == T), aes(y=mean.b.dev, x=log.lambda, colour="lambda.min"), size=2 ) +          
    geom_vline(xintercept = subset(df, lambda.min == T)$log.lambda, colour='grey') +
    annotate("text", x=subset(df, lambda.min == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
             label=paste(round(df[subset(df, lambda.min == T)$lambda.at,c('mean', 'sme')],3), collapse='\n+/-'))

  if (nrow(subset(df, lambda.1se == T)) > 0) {
    gp = gp + geom_point(data=subset(df, lambda.1se == T), aes(y=mean, x=log.lambda, colour="lambda.1se"), size=2 ) +
      geom_vline(xintercept = subset(df, lambda.1se == T)$log.lambda, colour='grey') +
      annotate("text", x=subset(df, lambda.1se == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
               label=paste(round(df[subset(df, lambda.1se == T)$lambda.at,c('mean', 'sme')],3), collapse='\n+/-'))

    gpD = gpD + geom_point(data=subset(df, lambda.1se == T), aes(y=mean.b.dev, x=log.lambda, colour="lambda.1se"), size=2 ) +
      geom_vline(xintercept = subset(df, lambda.1se == T)$log.lambda, colour='grey') +
      annotate("text", x=subset(df, lambda.1se == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
               label=paste(round(df[subset(df, lambda.1se == T)$lambda.at,c('mean', 'sme')],3), collapse='\n+/-')) 
  }

@@

```{r xvalpt, message=F, warning=F, echo=T, fig.height=16, fig.width=16}
nl = 1000

@@

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  for (a in alpha.values) {
    fit0 <- glmnet(dysplasia.df, labels, alpha=a, nlambda=nl, family='binomial') # all patients
    #autoplot(fit0, xvar='lambda', main=paste('fit0, all samples, alpha=',a,sep='')) + theme(legend.position='none')

    l = fit0$lambda
    if (a == 0) {
      l = sort(c(fit0$lambda, seq(exp(-5), exp(-10), -1e-6),
            seq(exp(-10), exp(-15), -1e-8),
            seq(exp(-15), exp(-20), -1e-9)), decreasing=T)
    }
    cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=l, pts=pts, a=a, nfolds=folds, splits=splits, fit=fit0)

@@

pg = dysplasia.df[ labels == 1 ,regions]

nonpg = dysplasia.df[ labels == 0 ,regions]

colnames(pg)

meantests = sapply(regions, function(r)   t.test(pg[,r], nonpg[,r])$p.value )

#which(meantests < 0.05)

summary(pg[,'1:119640299-129610322'])
summary(nonpg[,'1:119640299-129610322'])

@@

```{r, warning=F, message=F}
#topF = names(which((coef.stable[['1']]) >= 0.75))
topF = names(coef.stable[['1']])

features.in.prog = dysplasia.df[names(which(labels == 1)),topF]
features.nonprog = dysplasia.df[names(which(labels == 0)),topF]

feat = which(colnames(dysplasia.df) %in% topF)

# Gain or loss not correlated in prog or non

gl = cbind.data.frame(
  'P.gain' = apply(features.in.prog, 2, function(x) sum( as.integer(x >= 1.1) ) ), 
  'P.loss' = apply(features.in.prog, 2, function(x) sum( as.integer(x <= 0.9) ) ),
  'NP.gain' = apply(features.nonprog, 2, function(x) sum( as.integer(x >= 1.1) ) ),
  'NP.loss' = apply(features.nonprog, 2, function(x) sum( as.integer(x <= 0.9) ) ) )

allgl = cbind.data.frame(
  'ALL.P.gain' = apply(dysplasia.df[names(which(labels == 1)),], 2, function(x) sum( as.integer(x >= 1.1) ) ), 
  'ALL.P.loss' = apply(dysplasia.df[names(which(labels == 1)),], 2, function(x) sum( as.integer(x <= 0.9) ) ),
  'ALL.NP.gain' = apply(dysplasia.df[names(which(labels == 0)),], 2, function(x) sum( as.integer(x >= 1.1) ) ), 
  'ALL.NP.loss' = apply(dysplasia.df[names(which(labels == 0)),], 2, function(x) sum( as.integer(x <= 0.9) ) ) )

grid.arrange(
  ggcorr(gl, label=T) + ggtitle('Selected Features'),
  ggcorr(allgl, label=T) + ggtitle('All regions'), 
top='Gains vs Losses')

# Gains/losses enriched in the top features?
fisher.test(rbind(colSums(gl[,c('P.gain','P.loss')]),colSums(allgl[-feat,c('ALL.P.gain','ALL.P.loss')])  ))

fisher.test(rbind(colSums(gl[,c('NP.gain','NP.loss')]),colSums(allgl[-feat,c('ALL.NP.gain','ALL.NP.loss')])  ))

@@

#### bp Length

```{r, echo=F, warning=F, message=F}
chromInfo = read.table('hg19_genes.txt', sep='\t', header=T)
chromInfo$chr = factor(chromInfo$chr, levels=c(1:22, 'X','Y'))

topFSegments = do.call(rbind,  strsplit( topF, ':|-' ))
colnames(topFSegments) = c('chr','start','end')
topFSegments = as.data.frame(apply(topFSegments, 2, as.numeric))
topFSegments$chr = factor(topFSegments$chr, levels=chromInfo$chr)
topFSegments = arrange(topFSegments, chr, start)

topFRegions = arrange(merge(chromInfo[,c('chr','length', 'protein_coding')], topFSegments %>% group_by(chr) %>% summarise('totalbp'=sum(end-start), 'n.segments'=length(chr)), all.x=T), -protein_coding)

# Number of segments is strongly correlated with the length of the chromosome
with(topFRegions, cor.test(length, n.segments))

topFRegions$adj.info.content = topFRegions$protein_coding/(topFRegions$length^0.6)
with(topFRegions, cor.test(length, adj.info.content))

topFRegions = transform(topFRegions, ai.cut=cut(adj.info.content, 5, labels=c(5:1), include.lowest=T))           # bin data
ggplot(topFRegions, aes(reorder(chr, -adj.info.content), totalbp/length, fill=ai.cut)) + ylim(0, 0.5) +
  labs(y='Feature bp/total chr length', title='Ratio of covered base pairs', x='chr') +
  geom_bar(stat='identity') +  geom_text(aes(label=round(totalbp/length, 2)), vjust=-0.5) +
  scale_fill_discrete(name='Length adjusted information content') + theme(legend.position = 'bottom')

ct = with(topFRegions, cor.test(adj.info.content, totalbp/length))

bpCovered = with(topFRegions, sum(totalbp, na.rm=T)/sum(as.numeric(length)))
```

Strong correlation between the length adjusted information content (protein coding genes) and the covered bases in a chromosome r=`r round(ct$estimate, 2)` (p-value `r round(ct$p.value, 4)`). No features from chromosome 19 (the most gene dense chromosome) were selected.

`r round(bpCovered, 2)*100`% of the total bases are within the top selected features.

#### Genes?

Using Biomart get all genes that are fully within the the segment boundaries.

```{r, warning=F, message=F}
mart = biomaRt::useMart("ENSEMBL_MART_ENSEMBL", host="grch37.ensembl.org")
biomart = useDataset("hsapiens_gene_ensembl", mart)

attr = c('ensembl_gene_id','hgnc_symbol','start_position','end_position', 'chromosome_name', 'ensembl_exon_id','percentage_gene_gc_content')

bm = getBM(mart=biomart, attributes=attr, 
           filters=list('chromosomal_region'=sub('-', ':',topF), "biotype"=c("IG_C_gene","IG_D_gene","IG_J_gene","IG_V_gene","protein_coding","TR_C_gene","TR_D_gene","TR_J_gene","TR_V_gene")))
#head(bm)
bm = bm[-which(bm$hgnc_symbol == ""),]
bm = bm %>% group_by(hgnc_symbol, chromosome_name, start_position, end_position) %>% summarise('exon_count' = length(ensembl_exon_id)) 
exons = bm %>% group_by(chromosome_name) %>% summarise('exons'=sum(exon_count))

for (i in 1:nrow(topFSegments)) {
  s = topFSegments[i,]
  sbm = with(bm, which(chromosome_name == s[['chr']] & start_position >= s[['start']] & end_position <= s[['end']]))

  topFSegments[i,'genes'] = length(sbm)
  topFSegments[i,'exons'] = sum(bm[sbm,'exon_count'])

  bm[sbm, 'segment.index'] = i
}

# Not related to the length of the segment
with(topFSegments, cor.test(genes, (end-start)))

head(bm)

topFRegions = merge(topFRegions, (bm %>% group_by(chromosome_name) %>% summarise('gene.cnt'=length(unique(hgnc_symbol)), 'exon.cnt'=sum(exon_count))), by.x='chr', by.y='chromosome_name', all.x=T)

with(topFRegions, cor.test(protein_coding, exon.cnt))
with(topFRegions, cor.test(gene.cnt, exon.cnt)) # expected

# Exons correlate with the adjusted information content measure
with(topFRegions, cor.test(exon.cnt, adj.info.content))

ggplot(topFRegions, aes(reorder(chr, -adj.info.content), gene.cnt/protein_coding, fill=ai.cut)) + geom_bar(stat='identity') +
  labs(y='Gene Ratio', x='chr', title='Feature genes vs all protein coding genes') +
  geom_text(aes(label=round(gene.cnt/protein_coding, 2)), vjust=-0.5) +
  scale_fill_discrete(name='Length adjusted information content') + theme(legend.position = 'bottom')

# Merge COSMIC genes
ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
ccgenes = ccgenes[-grep('X,Y', ccgenes$Genome.Location),]

# This is going to strip out some genes that don't appear in ensembl
attr = c('ensembl_gene_id','hgnc_symbol','start_position','end_position', 'chromosome_name')
ccbm = getBM(mart=biomart, attributes=attr, filters=list('external_gene_name'=ccgenes$Gene.Symbol, 'chromosome_name'=c(1:22,'X','Y')))
ccbm = ccbm[-which(ccbm$hgnc_symbol == ""),]
ccbm$chromosome_name = factor(ccbm$chromosome_name, levels=chromInfo$chr)

topFRegions = merge(topFRegions, (ccbm %>% group_by(chromosome_name) %>% summarise('COSMIC.all'=length(unique(hgnc_symbol)))), by.x='chr',by.y='chromosome_name', all.x=T)

containedCC = subset(bm, hgnc_symbol %in% ccbm$hgnc_symbol)

topFRegions = merge(topFRegions, (containedCC %>% group_by(chromosome_name) %>% summarise('COSMIC.cont'=length(unique(hgnc_symbol)))), by.x='chr',by.y='chromosome_name', all.x=T)

ggplot(topFRegions, aes(reorder(chr, -adj.info.content), COSMIC.cont/COSMIC.all, fill=ai.cut)) + geom_bar(stat='identity') +
  labs(y='COSMIC Gene Ratio', x='chr', title='COSMIC genes: Feature vs all')+
  geom_text(aes(label=round(COSMIC.cont/COSMIC.all, 2)), vjust=-0.5) +
  scale_fill_discrete(name='Length adjusted information content') + theme(legend.position = 'bottom')
```

However, the `r round(bpCovered, 2)*100`% of bases appear to include `r round(length(unique(bm$hgnc_symbol))/sum(chromInfo$protein_coding), 2)*100`% (`r length(unique(bm$hgnc_symbol))`) of known protein-coding genes.  Similarly, `r round(length(unique(containedCC$hgnc_symbol))/length(ccgenes$Gene.Symbol),2)*100`% of the `r length(ccgenes$Gene.Symbol)` COSMIC genes are contained within these regions.

This appears to be a fairly large enrichment.

@@

file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  # No HGD/IMC
  samples = intersect(rownames(dysplasia.df), subset(info, Pathology %nin% c('HGD', 'IMC'))$Samplename)
  for (a in alpha.values) {
    # all patients
    fitNoHGD <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=nl) 
    cv.nohgd = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitNoHGD$lambda, pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fitNoHGD)

@@

## Randomize the labels

# Leave one out and evaluate predictions

@@

    # Fit generated on all samples, including HGD
    a = 1
    fitLOO <- glmnet(training, labels[train.rows], alpha=a, family='binomial', nlambda=nl) # all patients
    #l = sort(c(fitLOO$lambda, seq(exp(-5), exp(-10), -0.001),
    #      seq(exp(-10), exp(-15), -1e-7), 
    #      seq(exp(-15), exp(-20), -1e-9)), decreasing=T)
    l = fitLOO$lambda
    cv = crossvalidate.by.patient(x=training, y=labels[train.rows], lambda=l, 
                                  pts=subset(pts, Samplename %in% samples), a=a, nfolds=10, splits=5, fit=fitLOO)
    #plots[[pt]] = cv$plot + ggtitle(pt)

@@

back = lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], function(pt) {
  pt = pt %>% group_by(PID, Endoscopy.Year) %>% summarise(
                  'pred'=sum(as.integer(Prediction > 0.5)), 'samples'=length(PID), 
                  #'final.endo'= length(which(grepl('HGD|IMC', Pathology))) > 0,
                  'months.before.final'= (max(pt$Endoscopy.Year) - unique(Endoscopy.Year))*12 )
  arrange(pt, -Endoscopy.Year, PID)
})

endos = do.call(rbind, lapply(1:max(sapply(back,nrow)), function(i) {
  df = do.call(rbind, lapply(back, function(x) {
    if (i > nrow(x)) return(NA)
    cbind(x[i, c('Endoscopy.Year','pred','samples', 'months.before.final')], i)
  }))
  df$Patient = rownames(df)
  return(df)
}))

@@

```{r, echo=F, warning=F, message=F, fig.height=8, fig.width=8, eval=F}
plot.endo( subset(endos, Patient %in% subset(me, max.endo <= 2 & max.endo > 0)$Patient), min(endos$months.before.final), max(endos$months.before.final)  )
```

```{r, echo=F, warning=F, message=F, fig.height=4, fig.width=4, eval=F}
plot.endo( subset(endos, Patient %in% subset(me, max.endo == 0)$Patient),min(endos$months.before.final), max(endos$months.before.final) )
```

@@

@@

## Data Quality

### By Batch first

```{r, echo=T, warning=F, message=F, eval=F}
all = do.call(rbind, pg.samp)

prog = do.call(rbind, pg.samp[subset(sum.patient.data, Status == 'P')$Patient])
nonp = do.call(rbind, pg.samp[subset(sum.patient.data, Status == 'NP')$Patient])

## Similar numbers of each in these batches so how did we do?
tfPR = sapply( c('Exome_subcohort','Batch_1','Batch_2','Batch_3','Batch_4'), function(b) {
  pred = subset(prog, Batch.Name == b)$Prediction
  predB = subset(nonp, Batch.Name == b)$Prediction

  tpr = length(which(pred > 0.5))/length(pred)
  fpr = length(which(predB > 0.5))/length(predB)
  cbind.data.frame('TPR'=tpr, 'FPR'=fpr)
})

tfPR

## Exome_cohort fpr driven by the two patients that we predicted to be progressors in nearly every sample
length(which(pg.samp$AD0591$Prediction > 0.5))
length(which(pg.samp$AH0329$Prediction > 0.5))

## Except the exome cohort, the TPR rate appears to be right around 48-68%
## FPR is over 20% in Batch_4 though

unique(subset(nonp, Batch.Name == 'Batch_4')$Patient)

pg.samp$AHM0100$Prediction > 0.5
pg.samp$AHM0121$Prediction > 0.5
pg.samp$AD0644$Prediction > 0.5

cb = combn( as.character(unique(all$Batch.Name)), 2)

batch =  sapply(1:ncol(cb), function(i) {
  sapply(c('SampleSD','SampleMEAN', 'Total.Reads'), function(x) {
    tt = t.test(subset(all, Batch.Name == cb[1,i])[[x]], subset(all, Batch.Name == cb[2,i])[[x]])  
    tt$p.value
  })
}) 
```

```{r, echo=T, warning=F, message=F, eval=F}
smp<-function(df) {
  df[,c('Prediction', 'Prediction.Dev.Resid', 'SampleMEAN', 'SampleSD', 'Barretts.Cellularity', 'Total.Reads', 'p53.Status', 'PID')]
}

prog = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Patient], smp) )
nonp = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Patient], smp) )

GGally::ggcorr(data=prog, method=c('pairwise','pearson'), label=T) + labs(title='Pearson correlation for several QC measures')

# Compare poorly predicted to well predicted based on the sample SD measures
ttfunc<-function(col, df=NULL) {
  tt = t.test(subset(df, Prediction < 0.5)[[col]], subset(df, Prediction > 0.5)[[col]])
  cbind.data.frame('p.val'=tt$p.value, 'statistic'=tt$statistic)
}

@@

```{r, echo=T, warning=F, message=F, eval=F}
 pander(sapply(colnames(prog)[-(1:2)], ttfunc, df=prog), caption='Progressor t-test stats for QC measures between predictions probabilies  >&< 0.5 ')

 pander(sapply(colnames(nonp)[-(1:2)], ttfunc, df=nonp), caption='Non-progressor t-test stats for QC measures between predictions probabilies  >&< 0.5 ')

hist(prog$SampleSD)

subset(prog, SampleSD <= 0.025)
